---
title: "IvyDashboard"
author: "Tim Jacobs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
header-includes: \newcommand{\graph}[3]{ \raisebox{-#1mm}{\includegraphics[height=#2em]{#3}}}
---

```{r libraryOpts, echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")
library("knitr")
library("ggplot2")
library("gplots")
library("scales")
library("ggalt")
library("tidyquant")
library("reshape2")
library("zoo")

opts_chunk$set(warning = FALSE,
               message = FALSE,
               echo = FALSE)

# quandlKey <- "vQwvpzts4Y8Z8bt3nsjc"
# quandl_api_key(quandlKey)
```


```{r getEtfData}
today <- Sys.Date()

etfs <- data.frame(ETF = c("DBA", "DBC", "IFGL", "IYR", "TIP",
                           "AGG", "IEMG", "EDC", "IEFA", "IJR", "ITOT"),
                   Type = c("Agriculture Futures", "Commodities", "Intl Real Estate",
                            "US Real Estate", "TIPs Bonds", "US Aggregated Bonds",
                            "Emerging Markets", "3x Emerging Markets", "Foreign Total",
                            "S&P Small Cap",
                            "S&P Total"),
                   "Commission Free" = c("No", "No", "Yes", "No", "Yes",
                                         "Yes", "Yes", "No", "Yes", "Yes", "Yes"),
                   check.names = FALSE)

# ETF Data
agricultural <- tq_get("DBA", get = "stock.prices",
                      from = "2004-01-01", to = today)
agricultural$etf <- "DBA"
agricultural$type <- etfs[etfs$ETF == "DBA",]$Type
agricultural <- tq_mutate(agricultural, select = adjusted,
                          mutate_fun = rollapply, FUN = mean,
                          width = 200, align = "right",
                          col_rename = "avg200")

commodities <- tq_get("DBC", get = "stock.prices",
                      from = "2004-01-01", to = today)
commodities$etf <- "DBC"
commodities$type <- etfs[etfs$ETF == "DBC",]$Type
commodities <- tq_mutate(commodities, select = adjusted,
                         mutate_fun = rollapply, FUN = mean,
                         width = 200, align = "right",
                         col_rename = "avg200")

internationalRealEstate <- tq_get("IFGL", get = "stock.prices",
                      from = "2004-01-01", to = today)
internationalRealEstate$etf <- "IFGL"
internationalRealEstate$type <- etfs[etfs$ETF == "IFGL",]$Type
internationalRealEstate <- tq_mutate(internationalRealEstate, select = adjusted,
                                     mutate_fun = rollapply, FUN = mean,
                                     width = 200, align = "right",
                                     col_rename = "avg200")

usRealEstate <- tq_get("IYR", get = "stock.prices",
                      from = "2004-01-01", to = today)
usRealEstate$etf <- "IYR"
usRealEstate$type <- etfs[etfs$ETF == "IYR",]$Type
usRealEstate <- tq_mutate(usRealEstate, select = adjusted,
                          mutate_fun = rollapply, FUN = mean,
                          width = 200, align = "right",
                          col_rename = "avg200")

tipsBond <- tq_get("TIP", get = "stock.prices",
                      from = "2004-01-01", to = today)
tipsBond$etf <- "TIP"
tipsBond$type <- etfs[etfs$ETF == "TIP",]$Type
tipsBond <- tq_mutate(tipsBond, select = adjusted,
                      mutate_fun = rollapply, FUN = mean,
                      width = 200, align = "right",
                      col_rename = "avg200")

aggregatedBond <- tq_get("AGG", get = "stock.prices",
                      from = "2004-01-01", to = today)
aggregatedBond$etf <- "AGG"
aggregatedBond$type <- etfs[etfs$ETF == "AGG",]$Type
aggregatedBond <- tq_mutate(aggregatedBond, select = adjusted,
                            mutate_fun = rollapply, FUN = mean,
                            width = 200, align = "right",
                            col_rename = "avg200")

emergingMarkets <- tq_get("IEMG", get = "stock.prices",
                      from = "2004-01-01", to = today)
emergingMarkets$etf <- "IEMG"
emergingMarkets$type <- etfs[etfs$ETF == "IEMG",]$Type
emergingMarkets <- tq_mutate(emergingMarkets, select = adjusted,
                             mutate_fun = rollapply, FUN = mean,
                             width = 200, align = "right",
                             col_rename = "avg200")

emergingMarkets3x <- tq_get("EDC", get = "stock.prices",
                      from = "2004-01-01", to = today)
emergingMarkets3x$etf <- "EDC"
emergingMarkets3x$type <- etfs[etfs$ETF == "EDC",]$Type
emergingMarkets3x <- tq_mutate(emergingMarkets3x, select = adjusted,
                             mutate_fun = rollapply, FUN = mean,
                             width = 200, align = "right",
                             col_rename = "avg200")

foreignTotal <- tq_get("IEFA", get = "stock.prices",
                      from = "2004-01-01", to = today)
foreignTotal$etf <- "IEFA"
foreignTotal$type <- etfs[etfs$ETF == "IEFA",]$Type
foreignTotal <- tq_mutate(foreignTotal, select = adjusted,
                          mutate_fun = rollapply, FUN = mean,
                          width = 200, align = "right",
                          col_rename = "avg200")

sAndPSmallCap <- tq_get("IJR", get = "stock.prices",
                      from = "2004-01-01", to = today)
sAndPSmallCap$etf <- "IJR"
sAndPSmallCap$type <- etfs[etfs$ETF == "IJR",]$Type
sAndPSmallCap <- tq_mutate(sAndPSmallCap, select = adjusted,
                           mutate_fun = rollapply, FUN = mean,
                           width = 200, align = "right",
                           col_rename = "avg200")

sAndPTotal <- tq_get("ITOT", get = "stock.prices",
                      from = "2004-01-01", to = today)
sAndPTotal$etf <- "ITOT"
sAndPTotal$type <- etfs[etfs$ETF == "ITOT",]$Type
sAndPTotal <- tq_mutate(sAndPTotal, select = adjusted,
                        mutate_fun = rollapply, FUN = mean,
                        width = 200, align = "right",
                        col_rename = "avg200")

assetPriceHistory <- rbind(agricultural, commodities, internationalRealEstate,
                           usRealEstate, tipsBond, aggregatedBond, emergingMarkets,
                           emergingMarkets3x, foreignTotal, sAndPSmallCap, sAndPTotal)
# Asset Returns
assetMonthlyReturns <- assetPriceHistory %>%
    group_by(etf) %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn,
                 period     = "monthly")

assetPriceHistory$position <- "invested"
assetPriceHistory[which(assetPriceHistory$adjusted <
                          assetPriceHistory$avg200),]$position <- "cash"

# Add a Cash asset for when we are not invested
# Need dates. We could create a sequence but we don't have
# a good way to find all the dates the market is closed
# we'll just use the same dates as the most complete asset
# data frame
l <- sapply(unique(assetPriceHistory$etf), FUN = function(x){
  nrow(assetPriceHistory[assetPriceHistory$etf == x,])
})
cashdates <-
  assetPriceHistory[assetPriceHistory$etf == names(l)[which(l == max(l))][1],]$date
cash <- data.frame(date = cashdates,
                   open = 1,
                   high = 1,
                   low = 1,
                   close = 1,
                   volume = 1,
                   adjusted = 1,
                   etf = "CASH",
                   type = "Cash",
                   avg200 = 1)
rm(l, cashdates)

# Isolate records for each new month. This is when
# decisions and reallocations are made
assetPriceHistory$newMonth <- FALSE
r <- rle(as.POSIXlt(assetPriceHistory$date)$mon)
assetPriceHistory[cumsum(r[["lengths"]]) + 1,]$newMonth <- TRUE
assetPriceHistory <- assetPriceHistory[complete.cases(assetPriceHistory),]
decisions <- assetPriceHistory[assetPriceHistory$newMonth,]

# Calculate average absolute return from 3, 6, and 12 month returns
a <- lapply(unique(decisions$etf), FUN = function(x){decisions[decisions$etf == x,]})
names(a) <- unique(decisions$etf)
decisions <-
  do.call(rbind, lapply(a,
                        FUN = function(x){
                          data.frame(x,
                                     lag3 = c(rep(NA, times = 2), diff(x$adjusted, lag = 2)),
                                     lag6 = c(rep(NA, times = 5), diff(x$adjusted, lag = 5)),
                                     lag12 = c(rep(NA, times = 11), diff(x$adjusted, lag = 11))
                                    )
                          })
          )

decisions <- mutate(decisions,
                    absReturn3 = lag3/adjusted*100,
                    absReturn6 = lag6/adjusted*100,
                    absReturn12 = lag12/adjusted*100,
                    avgReturn = (absReturn3 + absReturn6 + absReturn12)/3)

# Isolate Buy and Sell dates
r <- rle(decisions$position)
buySell <- decisions[cumsum(r[["lengths"]]) + 1,]
buySell <- buySell[complete.cases(buySell),]
buySell$newMonth <- NULL
colnames(buySell)[colnames(buySell)=="position"] <- "action"
buySell[which(buySell$action == "cash"),]$action <- "Sell"
buySell[which(buySell$action == "invested"),]$action <- "Buy"

rm(a, r)
```


# Average of 3, 6, and 12 Month Absolute Returns

```{r plotAbsReturns, fig.asp=0.4}
absReturnPlotData <- decisions[decisions$date == max(decisions$date),]
absReturnPlotData$etf <-
  factor(absReturnPlotData$etf,
         levels = absReturnPlotData[order(absReturnPlotData$avgReturn,
                                          decreasing = TRUE),]$etf,
         ordered = TRUE)
top4 <- head(levels(absReturnPlotData$etf), 4)

ggplot(absReturnPlotData) +
  geom_bar(aes(x = etf, y = avgReturn, fill = position),
           stat = "identity") +
  theme_tq() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("red", "green")) +
  ylab("Average Total Return")
```


```{r makePlotDataframes}
# Only use recent history
daysOfHistory <- 300
firstDate <- Sys.Date() - daysOfHistory

# Create data frame for returns plots
a <- decisions
a$date <- as.yearmon(a$date)
a$key <- paste(a$date, a$etf)

returnsPlotData <- assetMonthlyReturns[assetMonthlyReturns$date >= firstDate, ]
returnsPlotData$date <- as.yearmon(returnsPlotData$date)
returnsPlotData$key <- paste(returnsPlotData$date, returnsPlotData$etf)
returnsPlotData$position <- a[match(returnsPlotData$key, a$key),]$position
returnsPlotData$etf <- factor(returnsPlotData$etf,
                              levels(absReturnPlotData$etf),
                              ordered = TRUE)

# Create data frame for price plots
pricePlotData <- assetPriceHistory[assetPriceHistory$date >= firstDate, ]
a <- rle(pricePlotData$position)
groups <- rep(1:length(a$lengths), times = a$lengths)
pricePlotData$group <- groups
pricePlotData$etf <- factor(pricePlotData$etf,
                            levels(absReturnPlotData$etf),
                            ordered = TRUE)

rm(a)


# plot just the top 4 performing asset classes
returnsPlotDataTop4 <- returnsPlotData[returnsPlotData$etf %in% top4,]
returnsPlotDataBonds <- returnsPlotData[returnsPlotData$etf %in% c("TIP", "AGG"),]

pricePlotDataTop4 <- pricePlotData[pricePlotData$etf %in% top4,]
pricePlotDataBonds <- pricePlotData[pricePlotData$etf %in% c("TIP", "AGG"),]
rm(a, groups)

```


# Price Relative to 200 Day Moving Average

```{r relativePricePlot}
ggplot(assetPriceHistory[assetPriceHistory$date >= firstDate,]) +
  geom_line(aes(x = date, y = adjusted/avg200)) +
  geom_hline(yintercept = 1, color = "red") +
  facet_wrap(~etf) +
  theme_tq() +
  ylab("Adjusted Price as a % of 200 Day Average") +
  scale_y_continuous(labels = percent)
```


# Bull Market Monthly Performance (top 4 highest returns)
Bar plots show the monthly returns. They are filled red if the monthly return is negative and green if positive. They are outlined red if the asset price is below the 200 day moving average on the monthly decision date, or green if the ending price is above the 200 day moving average.


```{r plotBullPerformance, fig.asp=0.65}
ggplot() +
  geom_line(data = pricePlotDataTop4,
            aes(x = date, y = adjusted/avg200-1)) +
  geom_bar(data = returnsPlotDataTop4,
           aes(x = as.Date(date), y = monthly.returns,
               fill = monthly.returns < 0,
               color = position),
           alpha = 0.4,
           stat = "identity") +
  facet_grid(etf~.) +
  theme_tq() +
  theme(legend.position = "none") +
  ylab("Bars represent monthly return\nlines are asset prices as percent of 200 day moving average") +
  scale_fill_manual(values = c("green", "red")) +
  scale_color_manual(values = c("green", "red")) 
```


# Bear Market Monthly Performance (Bonds)

```{r plotBearPerformance, fig.asp=0.65}
ggplot() +
  geom_line(data = pricePlotDataBonds,
            aes(x = date, y = adjusted/avg200-1)) +
  geom_bar(data = returnsPlotDataBonds,
           aes(x = as.Date(date), y = monthly.returns,
               fill = monthly.returns < 0,
               color = position),
           alpha = 0.4,
           stat = "identity") +
  facet_grid(etf~.) +
  theme_tq() +
  theme(legend.position = "none") +
  ylab("Bars represent monthly return\nlines are asset prices as percent of 200 day moving average") +
  scale_fill_manual(values = c("green", "red")) +
  scale_color_manual(values = c("red", "green")) 
```

